Mission-time LTL (MLTL) Formula Validation Via Regular Expressions\thanks{Work supported in part by NSF grant DMS-1950583 and NSF CAREER Award CNS-1552934.}

### 2023-06-16 paper/Paper.tex
updated paper
minor fixes: passive voice, grammar, typos

@@

\begin{document}
%
%\title{Contribution Title\thanks{Supported by organization x.}}
\title{Mission-time LTL (MLTL) Formula Validation Via Regular Expressions\thanks{Work supported in part by NSF grant DMS-1950583}}
%
\titlerunning{MLTL Formula Validation Via Regular Expressions}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%

@@

NASA, for example, has developed several tools that operate over temporal logic requirements, such as FRET\cite{GMRPSS20}, R2U2\cite{RS17}, and a PVS library \cite{CTGPD22} for the logic MLTL (Mission-time Linear Temporal Logic) \cite{RRS14,LVR22}. MLTL was the logic used recently in NASA's Robonaut2 verification project \cite{KZJZR20} and is being used currently for both design-time and runtime verification of the NASA Lunar Gateway Vehicle System Manager \cite{DBR21}. Other recent verification %verification-for-certification
efforts involving MLTL include a JAXA autonomous satellite \cite{JAXA}, a UAS Traffic Management (UTM) system involving Collins and Mosaic Aerospace \cite{HCHJR21}, a sounding rocket \cite{HLR21}, and multiple small satellites \cite{LLR21,LJBHCLR22,AJR22}. However, all of these successful verification efforts were carried out by groups specializing in formal methods research. To enable broader application of formal verification, and adoption across larger projects, we critically need better validation, e.g., so that analysis over MLTL-specified requirements can transparently contribute to flight certification.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \vspace{-0.1in}
\section{MLTL Regular Expressions} \label{regex}
% \vspace{-0.1in}
We modify the standard definition of Regular Expressions (Regex) to introduce notation that describes the satisfying computations of an MLTL formula. We begin by quoting the parts of the standard definition of a regex from \cite{sipser} that we use to describe our computations:

@@

For for regular expressions $w_0$ and $w_1$, a useful reduction is that $\{w_0 1 w_1, w_0 0 w_1\} = \{w_0 s w_1\}$. Each time set intersection is performed, we greedily apply this reduction to all appropriate pairs of regular expressions in the set. This prevents repeated set intersection operations from blowing up exponentially most of the time, and drastically improves running time.

@@

 $$S^{k} 1 S^{n-k-1}$$

@@

 $$S^{k} 0 S^{n-k-1}$$

>>>>

%% Save the class definition of \subparagraph
\let\llncssubparagraph\subparagraph
%% Provide a definition to \subparagraph to keep titlesec happy
\let\subparagraph\paragraph
%% Load titlesec
\usepackage[compact]{titlesec}
%% Revert \subparagraph to the llncs definition
\let\subparagraph\llncssubparagraph

@@

\begin{document}
%
%\title{Contribution Title\thanks{Supported by organization x.}}
\title{Mission-time LTL (MLTL) Formula Validation Via Regular Expressions\thanks{Work supported in part by NSF grant DMS-1950583 and NSF CAREER Award CNS-1552934.}}
%
\titlerunning{MLTL Formula Validation Via Regular Expressions}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%

@@

NASA, for example, has developed several tools that operate over temporal logic requirements, such as FRET\cite{GMRPSS20}, R2U2\cite{RS17}, and a PVS library \cite{CTGPD22} for the logic MLTL (Mission-time Linear Temporal Logic) \cite{RRS14,LVR22}. MLTL was the specification logic for NASA's Robonaut2 verification project \cite{KZJZR20} and is currently the specification logic for both design-time and runtime verification of the NASA Lunar Gateway Vehicle System Manager \cite{DBR21}. Other recent verification %verification-for-certification
efforts involving MLTL include a JAXA autonomous satellite \cite{JAXA}, a UAS Traffic Management (UTM) system involving Collins and Mosaic Aerospace \cite{HCHJR21}, a sounding rocket \cite{HLR21}, and multiple small satellites \cite{LLR21,LJBHCLR22,AJR22}. However, all of these successful verification efforts were carried out by groups specializing in formal methods research. To enable broader application of formal verification, and adoption across larger projects, we critically need better validation, e.g., so that analysis over MLTL-specified requirements can transparently contribute to flight certification.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \vspace{-0.1in}
\section{MLTL Regular Expressions} \label{regex}
% \vspace{-0.1in}
We modify the standard definition of Regular Expressions (regex) to introduce notation that describes the satisfying computations of an MLTL formula. We begin by quoting the parts of the standard definition of a regex from \cite{sipser} that we use to describe our computations:

@@

For for regular expressions $w_0$ and $w_1$, a useful reduction is that $\{w_0 1 w_1, w_0 0 w_1\} = \{w_0 s w_1\}$. Each time we perform set intersection, we greedily apply this reduction to all appropriate pairs of regular expressions in the set. This prevents repeated set intersection operations from blowing up exponentially most of the time, and drastically improves running time.

@@

 $$S^{k} 1 S^{n-k-1}$$,

@@

 $$S^{k} 0 S^{n-k-1}$$,
